### 2016-10-03 paper/algorithms_v2.tex
stashing paper edits (new intro outline, reorganized some section, clarified some notation)

\section{Goal-Oriented Inference Algorithm}\label{sect:alg}
%
Based on the theoretical developments in the last section, we now give a goal-oriented inference algorithm that allows one to combine models of varying fidelity, while maintaining rigorous control of QoI error.

Just as error estimates can be used to guide mesh-refinement~\cite{BecRann01}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide the division of the domain for a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. \Cref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMF}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, be a reduced-order model, or have a reduced parameter space~\cite{thesis/SIAM talk?}. The two models could also correspond to two levels of mesh-refinement, though in this case the method described in~\cite{BecVex05} could be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead. The derived error estimate is not applicable to all models, however. The two models have to be expressed in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. We need some degree of compatibility between the two models; namely, we assume that $\Psi_{LF}$ will be in a space admissible to $\mathcal{M}'_{HF,\Psi}$, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$.

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Extension to Offline-Online Setting}\label{sec:onoff}
%------------------------------------------------------------------------------------------------------------------------%

\Cref{alg:refSeries} can be naturally extended to an offline-online setting. One variation would be analagous to the offline-online decomposition proposed in \cite{LiebWill13}. The offline phase would consist of adaptively creating a mixed-fidelity model with an appropriate error tolerance given some expected observations $d^*$. When new data is received, one can then solve the inverse problem with the chosen mixed-fidelity model and the new data, and, if desired, compute an error estimate for the QoI. The mixed-fidelity inverse problems can be expected to require less time to solve than the high-fidelity inverse problems.

Another variation occurs in the context of a quasi-incremental data assimilation approach. A multi-fidelity model adaptive constraint can be generated for a given dataset $d_1$ in the offline phase. In the online phase, additional data points are included to form an augmented dataset $d_2$, and instead of solving the complete inverse problem with all the data points and the high-fidelity constraints, an iterative algorithm can be initiated using the multi-fidelity representation developed in the offline phase. The adaptive procedure, with the supplementary adjoint can be used to adapt the preexisting multi-fidelity representation to the new data points. In such a situation where $d_1\in d_2$, we expect that the model refinement will be concentrated around those of the new data points that also inform the QoI. Thus, in the online phase, even though the inverse problem will be solved with all data points $d_2$, the use of high-fidelity models will be limited, and the bulk of the computation (computing the auxilary variables, and supplementary adjoint) will be focused on identifying those new data points, which have the maximum impact on the QoI.

@@

We instead use the alternative method described in \cite{vanOpstaletal15}, decomposing the error estimate into contributions from locally supported basis functions rather than elements. For convenience of notation in this section, we drop subscripts so that $\Psi=\Psi_{LF}$, $Q=Q_{HF}$, and $U=U_{HF}$; for our numerical experiments, we have $Q_{LF}\subseteq Q$ and $U_{LF}=U$. Note that the error estimate in \cref{eq:finErrExp} can be equivalently written as,
%
\begin{equation}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi)(\Lambda)+\mathcal{L}'_{HF,\xi}(\xi_\Psi)(\chi_\Psi)+\mathcal{R}(e^3). \nonumber
\end{equation}
%
We consider a finite-dimensional \red{conforming} subspace $(Q^h\times U^h\times U^h)^3 \subset (Q\times U\times U)^3$ which contains the approximations $(\Lambda^h,\chi_\Psi^h)$. Define a basis $\Phi^h=\{(\varphi,\upvarphi)_i\}_{i\in I}$ consisting of locally supported functions such that span $\Phi^h=(Q^h\times U^h\times U^h)^3$; we can then write $(\Lambda^h,\chi_\Psi^h)=(\sum\limits_{i\in I}\varphi_i\lambda_i,\sum\limits_{i\in I}\upvarphi_i \chi_i)$. The error estimate,
%
\begin{equation}
\epsilon = -\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi)(\Lambda)+\mathcal{L}'_{HF,\xi}(\xi_\Psi)(\chi_\Psi) \leq \sum_{i\in I} \varepsilon_i,
\end{equation}
%
where,
%
\begin{equation}\label{eq:basisblame}
\varepsilon_i = \left| -\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi)(\lambda_i\varphi_i)+\mathcal{L}'_{HF,\xi}(\xi_\Psi)(x_i\upvarphi_i) \right|
\end{equation}
%
can be interpreted as the error contribution from the basis function $(\varphi,\upvarphi)_i$. Near the interfaces between the low-fidelity and high-fidelity regions, basis functions may have their support divided between the two regions and thus have a nonzero error contribution. The basis functions with the largest error contributions are flagged, and the elements in their support are refined.

%does B+V's method for 'cheaply' getting auxiliary variables not really work when parameters are a field and not handful of scalars? does it even matter, since bulk of costs seem to be going to super-adjoint?

%% Although Equation (\ref{eq:finErrExp}) is exact, the error estimate that can be calculated in practice will not generally be exact. Let us refer to a goal-oriented inverse problem as linear when the state $u$ and parameters $q$ are linearly related, the observation operator $C$ is linear in $u$, the regularization term $R$ is at most quadratic in the parameters, and the QoI functional $I$ is linear in $u$ and $q$. The remainder term $\mathcal{R}(e^3)$ is included in Equation (\ref{eq:finErrExp}) but would not, in practice, be calculated; in the case of a linear goal-oriented inverse problem, the remainder term disappears, but it is nonzero in general.

%% In motivating our approach, it is assumed that one can most accurately calculate the QoI from the parameter values inferred using the highest-fidelity forward model available, but that solving the inverse problem with this model is prohibitively expensive. It is also assumed that solving the inverse problem with a mixed-fidelity model, where this highest-fidelity model is only used in a portion of the domain, will be cheaper. There is a cost incurred by using our approach to design such a mixed-fidelity model, however, and it will sometimes be the case that the cost of obtaining this mixed-fidelity model exceeds that of just solving the inverse problem with the highest-fidelity model directly. Naively, if the auxiliary variables $\chi$ have $n$ degrees of freedom, they can be found by solving an $n\times n$ linear system, while the supplementary adjoint $\Lambda$ can be found by solving a $2n\times2n$ linear system. The cost of solving for the auxiliary variables can be reduced by using a technique described in \cite{BecVex05}, and the cost of solving for the supplementary adjoint $\Lambda$ can be reduced by reusing preconditioners. In general there are no guarantees that obtaining a mixed-fidelity model that meets the desired QoI error criterion will be less costly than just solving the inverse problem with the high-fidelity model. However, our approach targets problems for which solving the inverse problem with the high-fidelity model is prohibitively expensive, in which case it is expected that the cost of obtaining a satisfactory mixed-fidelity model will be comparatively low. Even in the case where a mixed-fidelity model for which the QoI error is adequately small cannot be found before another limit (for example, a maximum number of adaptive iterations) is reached, one still has an estimate for the error in the QoI without solving the prohibitively expensive inverse problem.

>>>>

\section{Goal-Oriented Inference Algorithm}\label{sec:alg}
%
Based on the theoretical developments in the last section, we now give a goal-oriented inference algorithm that allows one to combine models of varying fidelity, while maintaining rigorous control of QoI error.

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Adaptive Model Mixing ALgorithm}\label{sec:adapt_alg}
%------------------------------------------------------------------------------------------------------------------------%

Just as error estimates can be used to guide mesh-refinement~\cite{BecRann01,VendDarm00,PrudOden99,BecVex05,OTHERS}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide the division of the domain for a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. \Cref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMF}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, be a reduced-order model, or have a reduced parameter space. The two models could also correspond to two levels of mesh-refinement, though in this case the method described in~\cite{BecVex05} would likely be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead. The derived error estimate is not applicable to all models, however. The two models have to be expressed in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. We need some degree of compatibility between the two models; namely, we assume that $\Psi_{LF}$ will be in a space admissible to $\mathcal{M}'_{HF,\Psi}$, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$.

%------------------------------------------------------------------------------------------------------------------------%
%\subsection{Extension to Offline-Online Setting}\label{sec:onoff}
%------------------------------------------------------------------------------------------------------------------------%

%\Cref{alg:refSeries} can be naturally extended to an offline-online setting. One variation would be analagous to the offline-online decomposition proposed in \cite{LiebWill13}. The offline phase would consist of adaptively creating a mixed-fidelity model with an appropriate error tolerance given some expected observations $d^*$. When new data is received, one can then solve the inverse problem with the chosen mixed-fidelity model and the new data, and, if desired, compute an error estimate for the QoI. The mixed-fidelity inverse problems can be expected to require less time to solve than the high-fidelity inverse problems.

%Another variation occurs in the context of a quasi-incremental data assimilation approach. A multi-fidelity model adaptive constraint can be generated for a given dataset $d_1$ in the offline phase. In the online phase, additional data points are included to form an augmented dataset $d_2$, and instead of solving the complete inverse problem with all the data points and the high-fidelity constraints, an iterative algorithm can be initiated using the multi-fidelity representation developed in the offline phase. The adaptive procedure, with the supplementary adjoint can be used to adapt the preexisting multi-fidelity representation to the new data points. In such a situation where $d_1\in d_2$, we expect that the model refinement will be concentrated around those of the new data points that also inform the QoI. Thus, in the online phase, even though the inverse problem will be solved with all data points $d_2$, the use of high-fidelity models will be limited, and the bulk of the computation (computing the auxilary variables, and supplementary adjoint) will be focused on identifying those new data points, which have the maximum impact on the QoI.

@@

We instead use the alternative method described in \cite{vanOpstaletal15}, decomposing the error estimate into contributions from locally supported basis functions rather than elements. For ease of implementation, we can choose these basis functions to be the same as those spanning the space in which we solve for the primary, auxiliary, and supplementary adjoint vatriables. Near the interfaces between the low-fidelity and high-fidelity regions, basis functions may have their support divided between the two regions and thus have a nonzero error contribution. The basis functions with the largest error contributions are flagged, and the elements in their support are added to the high-fidelity subdomain. Basis functions whose support lie completely in the high-fidelity region will have a zero contribution and the elements in their support are never flagged for refinement again.

### 2016-10-03 paper/mathematical_formulation_v2.tex
renamed paper tex files so that we can do latexdiff on them
stashing paper edits (new intro outline, reorganized some section, clarified some notation)

\section{Mathematical Formulation}\label{sect:form}
%
We first introduce the goal oriented inverse problem formally, providing definitions and optimization, variational statments. Then we will derive a rigorous, a posteriori error estimate for the error induced in the QoI due to the use of mixed/lower fidelity constraints.

%------------------------------------------------------------------------------%
\subsection{Problem Setup}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
Let $Q$, $U$ be Hilbert spaces. We seek to infer an unknown parameter $q \in Q$. The parameter is constrained by a state equation, with a state variable $u \in U$, which satisfies a bilinear form $a:U \times U \to \Reals$, such that,
%
\begin{equation}
\label{eq:weakForm}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U.
\end{equation}
%
The bilinear form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses. Further, we define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations. The actual observations (data) are denoted by $d\in\R^{n_d}$.

The unknown parameter $q$ can then be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the constraints given by \cref{eq:weakForm}.

The constraints given by \cref{eq:weakForm} are typically models of physical processes or systems. However, a given physical system need not have a unique model that can describe it; there may be various models of different fidelities. This naturally introduces a hierarchy of models and, a tradeoff between fidelity and computational expense. Consider two models with which we infer parameters: a high-fidelity (HF) model and a low fidelity (LF) model. We then have a specific form of the inverse problem in \cref{eq:invOpt} with this model,
\begin{equation}
\begin{array}{r@{}c}
\min\limits_{q_{HF},u_{HF}} & \quad J_{HF}(q_{HF},u_{HF})=\frac{1}{2}\|d-C_{HF}(u_{HF})\|_2^2 + R_{HF}(q_{HF}) \\ \textrm{s.t. }& \quad a_{HF}(u_{HF},q_{HF})(\phi_{HF})=\ell_{HF}(q_{HF})(\phi_{HF}),\quad\forall\phi_{HF}\in U_{HF},
\end{array}
\end{equation}
for which we have the Lagrangian,
\begin{multline}
\mathcal{L}_{HF}(q_{HF},u_{HF},z_{HF}) = \\
J_{HF}(q_{HF},u_{HF})-(a_{HF}(u_{HF},q_{HF})(z_{HF})-\ell_{HF}(q_{HF})(z_{HF})).
\end{multline}
The inverse problem formulation and Lagrangian with the lower fidelity model can be similarly defined, and be identified with subscript $LF$.

In the case of a goal-oriented inverse problem, the ultimate purpose of inferring the unknown parameters is to calculate some Quantity of Interest (QoI). Assuming a single scalar QoI, we denote this QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI. Since our ultimate goal is to compute this QoI, the particular tradeoff we consider is the fidelity and resultant computational expense of the model we use, versus the error in the QoI. We thus seek to introduce the QoI functional $I$ into the inverse problem formulation, by introducing auxilary variables and additional adjoint equations. Then we use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those which occur due to the use of constraints other than those given by \cref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model. The main result is,
%
\begin{theorem}
\label{thm:error_estimate}
Following the notation established in \cref{sec:setup}, let the bilinear form $a:U \times U \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q\to\Reals$ be differentiable with respect to the parameter $q$, and the functional $I:Q\times U\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

@@

Denote by $\Psi_{HF}$ and $\Psi_{LF}$, stationary points of the high and low fidelity versions of \cref{eq:InvsOpt_auglag} and, consider the adjoint problem,
%
\begin{equation}
\label{eq:superAdjEq}
\mathcal{M}'_{HF,\Psi}(\Lambda;\Psi_{HF})(\Phi)=\mathcal{Q}(\Phi)=\mathcal M'_{HF}(\Psi_{LF})(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
\end{equation}
%
for the supplementary adjoint $\Lambda$. Then, the error in the Quantity of Interest $I$ is given by,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\end{multline}
%
where $\mathcal{R}$ is a remainder term that is third-order in the error $e=\Psi_{HF}-\Psi_{LF}$.
\end{theorem}
%
\begin{proof}
%
Observe that,
%
\begin{equation}
\label{eq:MeqI}
\mathcal{M}(\Psi)=I(q,u),
\end{equation}
%
since taking variations of $\mathcal{M}$ with respect to the auxiliary variables gives that $\xi_\Psi$ is a stationary point of $\mathcal{L}$.

Extending the property in \cref{eq:MeqI} to the augmented Lagrangians for the high and low fidelity models, we have,
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{,}
\end{equation}
Combining \cref{eq:repIwithM} and \cref{eq:beckvex} we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}

@@

Combining \cref{eq:preadj} and \cref{eq:adjOutErr}, we have,
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3). \nonumber
\end{multline}
%
which completes the proof.
\end{proof}
%
The above error estimate is general and the lower-fidelity model can also be a mixed-fidelity model that combines the high- and low-fidelity model. Given a low-fidelity model and a high-fidelity model, an intermediate, mixed-fidelity (MF) model can be formed by using the high-fidelity model in some parts of the domain, and the low-fidelity model in the rest of the domain.

>>>>

\section{Mathematical Formulation}\label{sec:form}
%
We first introduce the goal-oriented inverse problem formally. Then we derive a rigorous a posteriori error estimate for the error induced in the QoI due to the use of mixed/lower-fidelity models.

%------------------------------------------------------------------------------%
\subsection{Problem Setup}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
Consider a model for which the Galerkin formulation of the weak form is written as
\begin{equation}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\label{eq:weakForm}
\end{equation}
where $u\in U$ is the state, $q\in Q$ are the unknown parameters, $\phi$ is a test function, and the state space $U$ and parameter space $Q$ are Hilbert spaces. The form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses (in \cref{eq:weakForm}, they are linear with respect to $\phi$). Further, we define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations. The actual observations (data) are denoted by $d\in\R^{n_d}$.

The unknown parameter $q$ can be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the model given by \cref{eq:weakForm}, which appears as a constraint in \cref{eq:invOpt_cons}. The constraints \cref{eq:invOpt_cons} are typically models of physical processes or systems. A given physical system can be described to varying degrees of fidelity using different models. This naturally introduces a hierarchy of models and a tradeoff between fidelity and computational expense; how one manages this tradeoff depends on what aspect of the solution one is interested in.

In the case of a goal-oriented inverse problem, the ultimate purpose of inferring the unknown parameters is to calculate some Quantity of Interest (QoI). Assuming a single scalar QoI, we denote this QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI. Since our ultimate goal is to compute this QoI, the particular tradeoff we consider is in the fidelity and resultant computational expense of the model we use, versus the error in the QoI. We thus seek to introduce the QoI functional $I$ into the inverse problem formulation by introducing auxilary variables and additional adjoint equations. Then we use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those which occur due to the use of difference models in the constraint \cref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model. 
%
\begin{theorem}
\label{thm:error_estimate}
Consider the inverse problem described by the constrained optimization problem \cref{eq:invOpt}. Let the form $a:U \times U \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q\to\Reals$ be differentiable with respect to the parameter $q$, and the functional $I:Q\times U\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

@@

Consider two models with which parameters can be inferred: a high-fidelity model and a lower-fidelity model. Let the high-fidelity (HF) and low-fidelity (LF) models, and their corresponding variables, Lagrangians, and augmented Lagrangians, be distinguished by the subscripts $_{HF}$ and $_{LF}$, respectively. In particular, let $\Psi_{HF}$ and $\Psi_{LF}$ denote the stationary points of the high- and low- fidelity augmented Lagrangians $\mathcal{M}_{HF}$ and $\mathcal{M}_{LF}$, respectively. Consider the adjoint problem
%
\begin{equation}
\label{eq:superAdjEq}
\mathcal{M}'_{HF,\Psi}(\Lambda;\Psi_{HF})(\Phi)=\mathcal{Q}(\Phi)=\mathcal M'_{HF}(\Psi_{LF})(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
\end{equation}
%
for the supplementary adjoint $\Lambda$, where $\Phi$ is a test function. Then, the error in the Quantity of Interest $I$ is given by,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\end{multline}
%
where $\mathcal{R}$ is a remainder term that is third-order in the error $e=\Psi_{HF}-\Psi_{LF}$.
\end{theorem}
%
\begin{proof}
%
Observe that,
%
\begin{equation}
\label{eq:MeqI}
\mathcal{M}(\Psi)=I(q,u),
\end{equation}
%
since taking variations of $\mathcal{M}$ with respect to the auxiliary variables gives that $\xi_\Psi$ is a stationary point of $\mathcal{L}$.

Extending the property in \cref{eq:MeqI} to the augmented Lagrangians for the high and low fidelity models, we have,
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{equation}
Combining \cref{eq:repIwithM} and \cref{eq:beckvex} we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}

@@

Combining \cref{eq:preadj} and \cref{eq:adjOutErr}, we have,
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3), \nonumber
\end{multline}
%
which completes the proof.
\end{proof}
%
